---
title: "Multiple and Logistic Regression Presentation---Group#1"
author: "Fiona, Samantha, Amanda, Jefferson"
date: 'Last compiled: `r format(Sys.time(), "%b %d, %Y")`'
output: 
  ioslides_presentation:
    logo: "./PICS/UCI.png"
---

```{r setup, include=FALSE, warning = FALSE, message = FALSE}
library(tidyverse)
knitr::opts_chunk$set(echo = FALSE)
```


```{r, message=FALSE}
red <- read_delim("winequality-red.csv", delim = ';')
white <- read_delim("winequality-white.csv", delim = ';')

# Add red category to red data
temp_red <- red %>% 
  mutate(type = "red")

# Add white category to white data
temp_white <- white %>% 
  mutate(type = "white")

# Combine data into super wine data! 
wine <- bind_rows(temp_red, temp_white)

# Remove spaces and replace with underscores to be tidy
colnames(wine) <- gsub("\\s", "_", colnames(wine))
```

## Wine Quality Dataset
- The dataset we chose was the Wine Quality Dataset 
- This actually includes two different datasets, one for red wine and one for white wine 
- Variables of the red and white wine include fixed and volatile acidity, citric acid, residual sugar, chorides, total sulfur dioxide, free sulfur dioxide, density, pH, sulphates, alcohol, and quality
- The wines included are mostly average quality wines, with very few poor or excellent quality ones
- We combined the red and white wine dataset into one large compilation of the data


## Reading in Dataset

```
red <- read_delim("winequality-red.csv", delim = ';')
white <- read_delim("winequality-white.csv", delim = ';')

temp_red <- red %>% 
  mutate(winecolor = "red")
  
temp_white <- white %>% 
  mutate(winecolor = "white")
  
wine <- bind_rows(temp_red, temp_white)

colnames(wine) <- gsub("\\s", "_", colnames(wine))
```

##Creating Parallel Slopes Model

```
mod <- lm(quality ~ total_sulfur_dioxide + type, data = wine)
library(broom)
augmented_mod <- augment(mod)

plot <- ggplot(augmented_mod, aes(x = total_sulfur_dioxide, y = quality, color = type)) +
  geom_point() +
  geom_line(aes(x = total_sulfur_dioxide, y = .fitted)) +
  theme_bw()
plot
```

```{r, echo = FALSE}
lm(quality ~ total_sulfur_dioxide + type, data = wine)
```


##Parallel Slopes

```{r fig.dim=1, fig.cap= "", echo=FALSE}
library(png)
library(grid)
img <- readPNG("parallel_slopes.png")
grid.raster(img)
```

##Creating Logistic Model
```
modL <- glm(type ~ total_sulfur_dioxide, data = wine, family = binomial)
```

```{r fig.dim=.5, fig.cap= "", echo=FALSE}
img <- readPNG("log_results.png")
grid.raster(img)
```

##Logistic Graph
```
plotL <- ggplot(wine, aes(x = total_sulfur_dioxide, y = type)) + 
  geom_jitter(width = 0, height = 0.05, alpha = 0.5) +
  geom_smooth(method = "glm", se = FALSE, method.args = list(family = "binomial"))
```
```{r fig.dim=1, fig.cap= "", echo=FALSE}
img <- readPNG("log_graph.png")
grid.raster(img)
```

## Creating Multiple Linear Regression Model

```
mod2 <- lm(quality ~ alcohol + type, data = wine)
augmented_mod2 <- augment(mod2)
str(augmented_mod2)
plot2 <- ggplot(augmented_mod2, aes(x = alcohol, y = quality, color = type)) + 
  geom_smooth(se = TRUE, method = loess) + 
  geom_line(aes(x = alcohol, y = .fitted)) + 
  theme_light()

plot2
```
## Multiple Regression Model
```{r fig.dim=2, fig.cap= "", echo=FALSE}
multi <- readPNG("multi.png")
grid.raster(multi)
```

## Citation
P. Cortez, A. Cerdeira, F. Almeida, T. Matos and J. Reis.
Modeling wine preferences by data mining from physicochemical properties. In Decision Support Systems, Elsevier, 47(4):547-553, 2009.
