---
title: "Toledo Exam Analysis"
author: "Josh Arford"
date: "9/12/2019"
output: slidy_presentation
  
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r, include = FALSE}
library(readxl)
library(ggplot2)
library(tidyverse)
library(moderndive)
library(stringr)
library(dplyr)
toledo <- read_xlsx("2017 Toledo - All (1).xlsx")

```

#Clean boi

```{r}
names(toledo) <- (c("banner", "score", "gpa", "hours", "matclass", "matgrade","chemclass", "chemgrade", "major", "studenttype", "prevresult"))

toledo$matgrade <- factor(toledo$matgrade)
levels(toledo$matgrade)[1] <- 4.0
levels(toledo$matgrade)[2] <- 3.7
levels(toledo$matgrade)[3] <- 3.3
levels(toledo$matgrade)[4] <- 3.0
levels(toledo$matgrade)[5] <- 2.7
levels(toledo$matgrade)[6] <- 2.3
levels(toledo$matgrade)[7] <- 2.0
levels(toledo$matgrade)[8] <- 1.7
levels(toledo$matgrade)[9] <- 1.3
levels(toledo$matgrade)[10] <- 1.0
levels(toledo$matgrade)[11] <- 0.7
levels(toledo$matgrade)[12] <- 0.3
levels(toledo$matgrade)[13] <- 0.0

toledo$chemgrade <- factor(toledo$chemgrade)
levels(toledo$chemgrade)[1] <- 4.0
levels(toledo$chemgrade)[2] <- 3.7
levels(toledo$chemgrade)[3] <- 3.3
levels(toledo$chemgrade)[4] <- 3.0
levels(toledo$chemgrade)[5] <- 2.7
levels(toledo$chemgrade)[6] <- 2.3
levels(toledo$chemgrade)[7] <- 2.0
levels(toledo$chemgrade)[8] <- 1.7
levels(toledo$chemgrade)[9] <- 1.3
levels(toledo$chemgrade)[10] <- 1.0
levels(toledo$chemgrade)[11] <- 0.7
levels(toledo$chemgrade)[12] <- 0.3
levels(toledo$chemgrade)[13] <- 0.0

toledo1 <- toledo[-36, ]





pass <- filter(toledo1, score > 15)
fail <- filter(toledo1, score <= 15)
pass14 <- filter(toledo1, score > 13)
fail14 <- filter(toledo1, score > 13)


```






#Analysis1

```{r, echo = FALSE}
#Histogram of all Exam Scores
hist(toledo1$score, main = "Distribution of All Exam Scores", breaks = 10, border = "orange", col = "blue", xlab = "Score", labels = TRUE, ylim = c(0, 10) ) 
```

#Analysis2

```{r, fig.height=4, fig.width=5}
#Histogram of all Passing Exam Scores
hist(as.numeric(pass$score), main = "Distribution of Passing Exam Scores", breaks = 10, border = "orange", col = "blue", xlab = "Score", labels = TRUE ) 
```

#Analysis3

```{r, fig.height=4, fig.width=5}

#Histogram of Failing Exam Scores
hist(as.numeric(fail$score), main = "Distribution of GPA for Failing Exam Scores", breaks = 10, border = "orange", col = "blue", xlab = "Score", labels = TRUE, ylim = c(0, 20)) 
```

#Analysis 4

```{r, echo = FALSE}
ggplot(toledo1, aes(x = as.numeric(gpa), y = as.numeric(score))) + 
  geom_point() + geom_vline(xintercept = 16, color = "green") + 
  ylab("GPA") +  
  xlab("Score") +
  ggtitle(expression("Score vs GPA"), subtitle = "Toledo Exam") + 
  theme(plot.background = element_rect(fill = 'black', colour = 'black')) + 
  theme(
            plot.title = element_text(color="gold", size=18, face="bold", hjust = .5),
            plot.subtitle = element_text(color="gold", size=10, face="bold", hjust = .5),
            axis.title.x = element_text(color="gold", size=14, face="bold"),
            axis.title.y = element_text(color="gold", size=14, face="bold"),
            axis.text.x=element_text(colour="gold"),
            axis.text.y=element_text(colour="gold")
) +  scale_x_continuous(name="GPA", limits=c(0, 4)) + 
  geom_smooth(method = "lm", se = F) 


```

#Big Graph yeeeyeyeee

```{r}
ggplot(toledo1, aes(x = as.numeric(chemgrade), y = as.numeric(matgrade))) + 
  geom_point() + geom_vline(xintercept = 16, color = "green") +
  ylab("GPA") +
  ggtitle(expression("Score vs GPA"), subtitle = "Toledo Exam") +
  theme(plot.background = element_rect(fill = 'black', colour = 'black')) +
  theme(
            plot.title = element_text(color="gold", size=18, face="bold", hjust = .5),
            plot.subtitle = element_text(color="gold", size=10, face="bold", hjust = .5),
            axis.title.x = element_text(color="gold", size=14, face="bold"),
            axis.title.y = element_text(color="gold", size=14, face="bold"),
            axis.text.x=element_text(colour="gold"),
            axis.text.y=element_text(colour="gold")
) +  scale_y_continuous(name="GPA", limits=c(0, 4)) +
  geom_smooth(method = "lm", se = F)


```

#Makin a data

```{r}
matgrade <- c(2.7, 2, 1, 3, 3.3, 3.3, 1.3, 2.3, 2.3, 2.7, 2.7, 3.3, 2.7, 3.3, 2.7, 2, 2, 2.7, 2.7, 2.3, 2.3, 2.3, 3.7, 2.7, 2.7, 0, 3.7, 1, .3, 2)
chemgrade <- c(3.3, 3.3, 2.3, 1, 2, 3.7, 2.3, 2.3, 1.7, 1.7, 3.3, 3.7, 2.3, .7, 3.3, 4, 1.3, .7, 3.7, 3, 4, 2.7, 2, 3.3, 3.7, .3, 3, 1.7, 1, 3)
score <- c(18, 11, 13, 18, 15, 18, 13, 11, 12, 11, 10, 13, 11, 13, 16, 17, 16, 11, 16, 14, 15, 12, 14, 15, 14, 12, 10, 12, 11, 13)

matchem <- data.frame(matgrade, chemgrade, score)
```


#Mathgrade vs score

```{r}
ggplot(matchem, aes(x = matgrade, y = score)) + geom_point() + geom_smooth(method = lm)
```

#Chemgrade vs score

```{r}
ggplot(matchem, aes(x = chemgrade, y = score)) + geom_point() + geom_smooth(method = lm)
```





##Notes


What if analyis
Logistical Regression
Look at what it should be


SAT scores - must be 550 or greater
25th percentile - 540
75th percentile - 630

70 percent of people have an SAT score high enough to pass Toledo exam

ACT Scores - must be 22 or greater
25th percentile - 22
75th percentile - 26

75 percent of people have an ACT score high enough to pass Toledo exam




##What if Analysis - What if the cut - off was 14

```{r, fig.height=4, fig.width=5}
#Histogram of all Passing Exam Scores
hist(as.numeric(pass14$score), main = "Normal Distribution of Passing Exam Scores", breaks = 10, border = "orange", col = "blue", xlab = "Score", labels = TRUE ) 
```

```{r, fig.height=4, fig.width=5}

#Histogram of Failing Exam Scores
hist(as.numeric(fail14$score), main = "Distribution of GPA for Failing Exam Scores", breaks = 10, border = "orange", col = "blue", xlab = "Score", labels = TRUE, ylim = c(0, 20)) 
```


```{r}
#fail5 <- fail1[-c(1, 2, 5, 7, 8, 10, 11, 12, 13, 15, 16, 24, 26, 27, 29, 31, 34, 37, 38), ] 





```










