---
title: "CRT DATA ANALYSIS"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, comment = NA, message = FALSE, prompt = FALSE, fig.pos = 'H')
library(knitr)
library(kableExtra)
CRT <- read.csv("CRT.csv")
```

## STEP ONE: Preparing the DATA Files
- For the CRT data there needs to be 2 columns. Delete all Columns but CRT SCORE and FINAL_GRADE
- Save the CRT data as CRT.csv and place it in the same directory as this RMD. (Comma Delimited) 


```{r}
# Gets rid of empty columns and rows
CRT <- subset(CRT, CRT.SCORE != "NA")[,1:2]

# Changes the Letter Grades to Numbers
CRT$FINAL_GRADE <- factor(CRT$FINAL_GRADE)
levels(CRT$FINAL_GRADE)[1] <- 4.0
levels(CRT$FINAL_GRADE)[2] <- 3.7
levels(CRT$FINAL_GRADE)[3] <- 3.0
levels(CRT$FINAL_GRADE)[4] <- 2.7
levels(CRT$FINAL_GRADE)[5] <- 3.3
levels(CRT$FINAL_GRADE)[6] <- 2.0
levels(CRT$FINAL_GRADE)[7] <- 1.7
levels(CRT$FINAL_GRADE)[8] <- 2.3
levels(CRT$FINAL_GRADE)[9] <- 1.0
levels(CRT$FINAL_GRADE)[10] <- 0.7
levels(CRT$FINAL_GRADE)[11] <- 1.3
levels(CRT$FINAL_GRADE)[12] <- 0.0

# Changes factors to numeric doubles
CRT$FINAL_GRADE <- as.numeric(levels(CRT$FINAL_GRADE)[CRT$FINAL_GRADE])

# Subsets the CRT data to the cutoff score of 11. CUTOFF Score can be adjusted here
CRT11 <- subset(CRT, CRT.SCORE >= 11)

# Subsets the CRT into passing and failing subsets
CRTNC <- subset(CRT11, FINAL_GRADE < 1.7)
CRTC <- subset(CRT11, FINAL_GRADE >= 1.7)

# Identifies proportions of fails observed
propCRTF <- round(length(CRTNC$FINAL_GRADE)/length(CRT$FINAL_GRADE),5)

# Plot of the grades based on scores
mod1 <- lm(CRT11$FINAL_GRADE~CRT11$CRT.SCORE)
plot(CRT11$CRT.SCORE,CRT11$FINAL_GRADE, col = "blue", pch = 20, xlab = "Score", ylab = "Grade")
abline(mod1, col = "red")
```


## Hypothesis

The null hypothesis ($H_0$) is that the failure rate of the CRT is 10%. The alternate hypothesis ($H_A$) is that the failure rate is greater than 10%. We get the following:

$$ H_0 = .10$$
$$ H_A > .10$$

## Test Statistic

The test statistic is the observed value that is being tested. In this case the test statistic is the `r round(propCRTF, 3)*100` % of student failures. 


```{r}
# Finds Standaard Deviation
CRT11sd <- sqrt((0.1*0.9)/length(CRT11$CRT.SCORE))

z11 <- (propCRTF - 0.1)/CRT11sd

# Finds p-value
p <- pnorm(z11, lower.tail = FALSE)
p
```

## Criteria

```{r}
# select alpha value here
alpha <- 0.1


# determines which statement to used based on p-value and alpha
if (p < alpha)
{
  A <- "Since the p-value is less than the alpha value so we reject the null hypothesis."
}else
{
  A <- "Since the p-value is not less than the alpha value so we fail to reject the null hypothesis"
}

```

Since the p-value is `r p` and the alpha value is `r alpha`. `r A`.


# Conclusion

```{r}
if (p < alpha)
{
  C<- "We find significant evidence to suggest that the failure rate is greater than 10%."
}else
{
  C<- "We fail to find evidence to suggest that the failure rate is greater than 10%."
}
```

`r C`

# Threshold and Conclusion of the Calculus Readiness Test

```{r}
#subsets all of the scores between 1 and 20

CRT1 <- subset(CRT, CRT.SCORE >= 1)
CRT2 <- subset(CRT, CRT.SCORE >= 2)
CRT3 <- subset(CRT, CRT.SCORE >= 3)
CRT4 <- subset(CRT, CRT.SCORE >= 4)
CRT5 <- subset(CRT, CRT.SCORE >= 5)
CRT6 <- subset(CRT, CRT.SCORE >= 6)
CRT7 <- subset(CRT, CRT.SCORE >= 7)
CRT8 <- subset(CRT, CRT.SCORE >= 8)
CRT9 <- subset(CRT, CRT.SCORE >= 9)
CRT10 <- subset(CRT, CRT.SCORE >= 10)
CRT11 <- subset(CRT, CRT.SCORE >= 11)
CRT12 <- subset(CRT, CRT.SCORE >= 12)
CRT13 <- subset(CRT, CRT.SCORE >= 13)
CRT14 <- subset(CRT, CRT.SCORE >= 14)
CRT15 <- subset(CRT, CRT.SCORE >= 15)
CRT16 <- subset(CRT, CRT.SCORE >= 16)
CRT17 <- subset(CRT, CRT.SCORE >= 17)
CRT18 <- subset(CRT, CRT.SCORE >= 18)
CRT19 <- subset(CRT, CRT.SCORE >= 19)
CRT20 <- subset(CRT, CRT.SCORE >= 20)

```

```{r}
# Number that failed
CRTF1NC <- subset(CRT1, FINAL_GRADE < 1.7)
CRTF2NC <- subset(CRT2, FINAL_GRADE < 1.7)
CRTF3NC <- subset(CRT3, FINAL_GRADE < 1.7)
CRTF4NC <- subset(CRT4, FINAL_GRADE < 1.7)
CRTF5NC <- subset(CRT5, FINAL_GRADE < 1.7)
CRTF6NC <- subset(CRT6, FINAL_GRADE < 1.7)
CRTF7NC <- subset(CRT7, FINAL_GRADE < 1.7)
CRTF8NC <- subset(CRT8, FINAL_GRADE < 1.7)
CRTF9NC <- subset(CRT9, FINAL_GRADE < 1.7)
CRTF10NC <- subset(CRT10, FINAL_GRADE < 1.7)
CRTF11NC <- subset(CRT11, FINAL_GRADE < 1.7)
CRTF12NC <- subset(CRT12, FINAL_GRADE < 1.7)
CRTF13NC <- subset(CRT13, FINAL_GRADE < 1.7)
CRTF14NC <- subset(CRT14, FINAL_GRADE < 1.7)
CRTF15NC <- subset(CRT15, FINAL_GRADE < 1.7)
CRTF16NC <- subset(CRT16, FINAL_GRADE < 1.7)
CRTF17NC <- subset(CRT17, FINAL_GRADE < 1.7)
CRTF18NC <- subset(CRT18, FINAL_GRADE < 1.7)
CRTF19NC <- subset(CRT19, FINAL_GRADE < 1.7)
CRTF20NC <- subset(CRT20, FINAL_GRADE < 1.7)

```

```{r}

PropCRT1NC <- round((length(CRTF1NC$FINAL_GRADE)/length(CRT$FINAL_GRADE)),5)
PropCRT2NC <- round((length(CRTF2NC$FINAL_GRADE)/length(CRT$FINAL_GRADE)),5)
PropCRT3NC <- round((length(CRTF3NC$FINAL_GRADE)/length(CRT$FINAL_GRADE)),5)
PropCRT4NC <- round((length(CRTF4NC$FINAL_GRADE)/length(CRT$FINAL_GRADE)),5)
PropCRT5NC <- round((length(CRTF5NC$FINAL_GRADE)/length(CRT$FINAL_GRADE)),5)
PropCRT6NC <- round((length(CRTF6NC$FINAL_GRADE)/length(CRT$FINAL_GRADE)),5)
PropCRT7NC <- round((length(CRTF7NC$FINAL_GRADE)/length(CRT$FINAL_GRADE)),5)
PropCRT8NC <- round((length(CRTF8NC$FINAL_GRADE)/length(CRT$FINAL_GRADE)),5)
PropCRT9NC <- round((length(CRTF9NC$FINAL_GRADE)/length(CRT$FINAL_GRADE)),5)
PropCRT10NC <- round((length(CRTF10NC$FINAL_GRADE)/length(CRT$FINAL_GRADE)),5)
PropCRT11NC <- round((length(CRTF11NC$FINAL_GRADE)/length(CRT$FINAL_GRADE)),5)
PropCRT12NC <- round((length(CRTF12NC$FINAL_GRADE)/length(CRT$FINAL_GRADE)),5)
PropCRT13NC <- round((length(CRTF13NC$FINAL_GRADE)/length(CRT$FINAL_GRADE)),5)
PropCRT14NC <- round((length(CRTF14NC$FINAL_GRADE)/length(CRT$FINAL_GRADE)),5)
PropCRT15NC <- round((length(CRTF15NC$FINAL_GRADE)/length(CRT$FINAL_GRADE)),5)
PropCRT16NC <- round((length(CRTF16NC$FINAL_GRADE)/length(CRT$FINAL_GRADE)),5)
PropCRT17NC <- round((length(CRTF17NC$FINAL_GRADE)/length(CRT$FINAL_GRADE)),5)
PropCRT18NC <- round((length(CRTF18NC$FINAL_GRADE)/length(CRT$FINAL_GRADE)),5)
PropCRT19NC <- round((length(CRTF19NC$FINAL_GRADE)/length(CRT$FINAL_GRADE)),5)
PropCRT20NC <- round((length(CRTF20NC$FINAL_GRADE)/length(CRT$FINAL_GRADE)),10)

```

```{r}

SD1 <- sqrt(0.1*0.9/length(CRT1$FINAL_GRADE))
SD2 <- sqrt(0.1*0.9/length(CRT2$FINAL_GRADE))
SD3 <- sqrt(0.1*0.9/length(CRT3$FINAL_GRADE))
SD4 <- sqrt(0.1*0.9/length(CRT4$FINAL_GRADE))
SD5 <- sqrt(0.1*0.9/length(CRT5$FINAL_GRADE))
SD6 <- sqrt(0.1*0.9/length(CRT6$FINAL_GRADE))
SD7 <- sqrt(0.1*0.9/length(CRT7$FINAL_GRADE))
SD8 <- sqrt(0.1*0.9/length(CRT8$FINAL_GRADE))
SD9 <- sqrt(0.1*0.9/length(CRT9$FINAL_GRADE))
SD10 <- sqrt(0.1*0.9/length(CRT10$FINAL_GRADE))
SD11 <- sqrt(0.1*0.9/length(CRT11$FINAL_GRADE))
SD12 <- sqrt(0.1*0.9/length(CRT12$FINAL_GRADE))
SD13 <- sqrt(0.1*0.9/length(CRT13$FINAL_GRADE))
SD14 <- sqrt(0.1*0.9/length(CRT14$FINAL_GRADE))
SD15 <- sqrt(0.1*0.9/length(CRT15$FINAL_GRADE))
SD16 <- sqrt(0.1*0.9/length(CRT16$FINAL_GRADE))
SD17 <- sqrt(0.1*0.9/length(CRT17$FINAL_GRADE))
SD18 <- sqrt(0.1*0.9/length(CRT18$FINAL_GRADE))
SD19 <- sqrt(0.1*0.9/length(CRT19$FINAL_GRADE))
SD20 <- sqrt(0.1*0.9/length(CRT20$FINAL_GRADE))

z1 <- (PropCRT1NC - 0.1)/SD1
z2 <- (PropCRT2NC - 0.1)/SD2
z3 <- (PropCRT3NC - 0.1)/SD3
z4 <- (PropCRT4NC - 0.1)/SD4
z5 <- (PropCRT5NC - 0.1)/SD5
z6 <- (PropCRT6NC - 0.1)/SD6
z7 <- (PropCRT7NC - 0.1)/SD7
z8 <- (PropCRT8NC - 0.1)/SD8
z9 <- (PropCRT9NC - 0.1)/SD9
z10 <- (PropCRT10NC - 0.1)/SD10
z11 <- (PropCRT11NC - 0.1)/SD11
z12 <- (PropCRT12NC - 0.1)/SD12
z13 <- (PropCRT13NC - 0.1)/SD13
z14 <- (PropCRT14NC - 0.1)/SD14
z15 <- (PropCRT15NC - 0.1)/SD15
z16 <- (PropCRT16NC - 0.1)/SD16
z17 <- (PropCRT17NC - 0.1)/SD17
z18 <- (PropCRT18NC - 0.1)/SD18
z19 <- (PropCRT19NC - 0.1)/SD19
z20 <- (PropCRT20NC - 0.1)/SD20

P1 <- pnorm(z1, lower.tail = FALSE)
P2 <- pnorm(z2, lower.tail = FALSE)
P3 <- pnorm(z3, lower.tail = FALSE)
P4 <- pnorm(z4, lower.tail = FALSE)
P5 <- pnorm(z5, lower.tail = FALSE)
P6 <- pnorm(z6, lower.tail = FALSE)
P7 <- pnorm(z7, lower.tail = FALSE)
P8 <- pnorm(z8, lower.tail = FALSE)
P9 <- pnorm(z9, lower.tail = FALSE)
P10 <- pnorm(z10, lower.tail = FALSE)
P11 <- pnorm(z11, lower.tail = FALSE)
P12 <- pnorm(z12, lower.tail = FALSE)
P13 <- pnorm(z13, lower.tail = FALSE)
P14 <- pnorm(z14, lower.tail = FALSE)
P15 <- pnorm(z15, lower.tail = FALSE)
P16 <- pnorm(z16, lower.tail = FALSE)
P17 <- pnorm(z17, lower.tail = FALSE)
P18 <- pnorm(z18, lower.tail = FALSE)
P19 <- pnorm(z19, lower.tail = FALSE)
P20 <- pnorm(z20, lower.tail = FALSE)
Score3 <- c(1:10)
Score4 <- c(11:20)
Prop.Fail3 <- c(PropCRT1NC,PropCRT2NC,PropCRT3NC,PropCRT4NC,PropCRT5NC,PropCRT6NC,PropCRT7NC,PropCRT8NC,PropCRT9NC, PropCRT10NC)
Prop.Fail4 <- c(PropCRT11NC,PropCRT12NC,PropCRT13NC,PropCRT14NC,PropCRT15NC,PropCRT16NC,PropCRT17NC,PropCRT18NC,PropCRT19NC, PropCRT20NC)
P.value3 <- c(P1, P2, P3, P4,P5,P6, P7,P8,P9,P10)
P.value4 <- c(P11, P12, P13, P14,P15,P16, P17,P18,P19,P20)
l <- cbind(Score3, Prop.Fail3,P.value3,Score4, Prop.Fail4,P.value4)
colnames(l) <- c("Score","Prop Fail", "p-value","Score","Prop Fail", "p-value")

kable(l) %>%
kable_styling(bootstrap_options = "striped", full_width = F, position = "center")



```





