---
title: "Pitching Data"
author: "Jefferson Ong and Josh Arford"
date: "9/10/2019"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```



```{r, include = FALSE}
library(readr)
library(tidyverse)
library(broom)
library(ggrepel)
library(moderndive)

vandy <- read.csv("vandy_1.csv")
vandy2 <- read.csv("vandy_2.csv")
vandy3 <- read.csv("vandy_3.csv")
```



```{r, include=FALSE}
## K means to determine pitch types
pitch_data <- read.csv("pitch_types.csv")
set.seed(1)
pitch_cluster <- kmeans(pitch_data[,3:6],2,nstart = 10)
pitch_cluster$centers
table(pitch_cluster$cluster, pitch_data$AutoPitchType)## K means to determine pitch types

library(cluster)
##png(file="kmeansPlot.png", height=500, width=650)
clusplot(pitch_data[,c(3:8)], pitch_cluster$cluster, color=T, lwd=2)
# dev.off()
```



```{r, include=FALSE}
# list(vandy)
```

# Pitcher 1

# RelSpeed, SpinRate, and SpinAxis with clusters of 2,3, and 4.

```{r}
## K means to determine pitch types

set.seed(1)
vandy_cluster <- kmeans(vandy[,8:10],2,nstart = 20)
vandy_cluster
table(vandy_cluster$cluster, vandy$AutoPitchType)## K means to determine pitch types

##png(file="kmeansPlot.png", height=500, width=650)
clusplot(vandy[,c(3:8)], vandy_cluster$cluster, color=T, lwd=2)
# dev.off()
```

```{r}
## K means to determine pitch types

set.seed(1)
vandy_cluster <- kmeans(vandy[,8:10],3,nstart = 20)
vandy_cluster
table(vandy_cluster$cluster, vandy$AutoPitchType)## K means to determine pitch types

##png(file="kmeansPlot.png", height=500, width=650)
clusplot(vandy[,c(3:8)], vandy_cluster$cluster, color=T, lwd=2)
# dev.off()
```

```{r}
## K means to determine pitch types

set.seed(1)
vandy_cluster <- kmeans(vandy[,8:10],4,nstart = 20)
vandy_cluster
table(vandy_cluster$cluster, vandy$AutoPitchType)## K means to determine pitch types

##png(file="kmeansPlot.png", height=500, width=650)
clusplot(vandy[,c(3:8)], vandy_cluster$cluster, color=T, lwd=2)
# dev.off()
```

# RelHeight and RelSide with clusters of 2,3, and 4

```{r}
## K means to determine pitch types

set.seed(1)
vandy_cluster <- kmeans(vandy[,12:13],2,nstart = 20)
vandy_cluster
table(vandy_cluster$cluster, vandy$AutoPitchType)## K means to determine pitch types

##png(file="kmeansPlot.png", height=500, width=650)
clusplot(vandy[,c(3:8)], vandy_cluster$cluster, color=T, lwd=2)
# dev.off()
```

```{r}
## K means to determine pitch types

set.seed(1)
vandy_cluster <- kmeans(vandy[,12:13],3,nstart = 20)
vandy_cluster
table(vandy_cluster$cluster, vandy$AutoPitchType)## K means to determine pitch types

##png(file="kmeansPlot.png", height=500, width=650)
clusplot(vandy[,c(3:8)], vandy_cluster$cluster, color=T, lwd=2)
# dev.off()
```

```{r}
## K means to determine pitch types

set.seed(1)
vandy_cluster <- kmeans(vandy[,12:13],4,nstart = 20)
vandy_cluster
table(vandy_cluster$cluster, vandy$AutoPitchType)## K means to determine pitch types

##png(file="kmeansPlot.png", height=500, width=650)
clusplot(vandy[,c(3:8)], vandy_cluster$cluster, color=T, lwd=2)
# dev.off()
```


# InducedVertBreak and HorzBreak with clusters of 2,3, and 4.


```{r}
## K means to determine pitch types

set.seed(1)
vandy_cluster <- kmeans(vandy[,16:17],2,nstart = 20)
vandy_cluster
table(vandy_cluster$cluster, vandy$AutoPitchType)## K means to determine pitch types

##png(file="kmeansPlot.png", height=500, width=650)
clusplot(vandy[,c(3:8)], vandy_cluster$cluster, color=T, lwd=2)
# dev.off()
```

```{r}
## K means to determine pitch types

set.seed(1)
vandy_cluster <- kmeans(vandy[,16:17],3,nstart = 20)
vandy_cluster
table(vandy_cluster$cluster, vandy$AutoPitchType)## K means to determine pitch types

##png(file="kmeansPlot.png", height=500, width=650)
clusplot(vandy[,c(3:8)], vandy_cluster$cluster, color=T, lwd=2)
# dev.off()
```

```{r}
## K means to determine pitch types

set.seed(1)
vandy_cluster <- kmeans(vandy[,16:17],4,nstart = 20)
vandy_cluster
table(vandy_cluster$cluster, vandy$AutoPitchType)## K means to determine pitch types

##png(file="kmeansPlot.png", height=500, width=650)
clusplot(vandy[,c(3:8)], vandy_cluster$cluster, color=T, lwd=2)
# dev.off()
```



# PlateLocHeight and PlateLocSide with clusters of 2,3, and 4

```{r}
## K means to determine pitch types

set.seed(1)
vandy_cluster <- kmeans(vandy[,18:19],2,nstart = 20)
vandy_cluster
table(vandy_cluster$cluster, vandy$AutoPitchType)## K means to determine pitch types

##png(file="kmeansPlot.png", height=500, width=650)
clusplot(vandy[,c(3:8)], vandy_cluster$cluster, color=T, lwd=2)
# dev.off()
```

```{r}
## K means to determine pitch types

set.seed(1)
vandy_cluster <- kmeans(vandy[,18:19],3,nstart = 20)
vandy_cluster
table(vandy_cluster$cluster, vandy$AutoPitchType)## K means to determine pitch types

##png(file="kmeansPlot.png", height=500, width=650)
clusplot(vandy[,c(3:8)], vandy_cluster$cluster, color=T, lwd=2)
# dev.off()
```

```{r}
## K means to determine pitch types

set.seed(1)
vandy_cluster <- kmeans(vandy[,18:19],4,nstart = 20)
vandy_cluster
table(vandy_cluster$cluster, vandy$AutoPitchType)## K means to determine pitch types

##png(file="kmeansPlot.png", height=500, width=650)
clusplot(vandy[,c(3:8)], vandy_cluster$cluster, color=T, lwd=2)
# dev.off()
```


# Pitcher 2 


```{r, include=FALSE}

library(dplyr)
vandy_2 <- select(vandy2, RelSpeed, SpinRate, SpinAxis, RelHeight, RelSide, InducedVertBreak, HorzBreak, PlateLocHeight, PlateLocSide)
vand2 <- na.omit(vandy_2)
```


# RelHeight and RelSide with clusters of 2,3, and 4

```{r}
set.seed(1)
cl <- kmeans(vand2[,1:3], 2, nstart = 10)


##png(file="kmeansPlot.png", height=500, width=650)
clusplot(vand2[,c(1:3)], cl$cluster, color=T, lwd=2)
```


```{r}
set.seed(1)
cl <- kmeans(vand2[,1:3], 3, nstart = 10)


##png(file="kmeansPlot.png", height=500, width=650)
clusplot(vand2[,c(1:3)], cl$cluster, color=T, lwd=2)
```


```{r}
set.seed(1)
cl <- kmeans(vand2[,1:3], 4, nstart = 10)


##png(file="kmeansPlot.png", height=500, width=650)
clusplot(vand2[,c(1:3)], cl$cluster, color=T, lwd=2)
```



# RelHeight and RelSide with clusters of 2,3, and 4

```{r}
set.seed(1)
cl <- kmeans(vand2[,4:5], 2, nstart = 10)


##png(file="kmeansPlot.png", height=500, width=650)
clusplot(vand2[,c(1:3)], cl$cluster, color=T, lwd=2)
```


```{r}
set.seed(1)
cl <- kmeans(vand2[,4:5], 3, nstart = 10)


##png(file="kmeansPlot.png", height=500, width=650)
clusplot(vand2[,c(1:3)], cl$cluster, color=T, lwd=2)
```






```{r}
set.seed(1)
cl <- kmeans(vand2[,4:5], 4, nstart = 10)


##png(file="kmeansPlot.png", height=500, width=650)
clusplot(vand2[,c(1:3)], cl$cluster, color=T, lwd=2)
```














# InducedVertBreak and HorzBreak with clusters of 2,3, and 4.


```{r}
set.seed(1)
cl <- kmeans(vand2[,6:7], 2, nstart = 10)


##png(file="kmeansPlot.png", height=500, width=650)
clusplot(vand2[,c(1:3)], cl$cluster, color=T, lwd=2)
```

```{r}
set.seed(1)
cl <- kmeans(vand2[,6:7], 3, nstart = 10)


##png(file="kmeansPlot.png", height=500, width=650)
clusplot(vand2[,c(1:3)], cl$cluster, color=T, lwd=2)
```

```{r}
set.seed(1)
cl <- kmeans(vand2[,6:7], 4, nstart = 10)


##png(file="kmeansPlot.png", height=500, width=650)
clusplot(vand2[,c(1:3)], cl$cluster, color=T, lwd=2)
```


# PlateLocHeight and PlateLocSide with clusters of 2,3, and 4


```{r}
set.seed(1)
cl <- kmeans(vand2[,8:9], 2, nstart = 10)


##png(file="kmeansPlot.png", height=500, width=650)
clusplot(vand2[,c(1:3)], cl$cluster, color=T, lwd=2)
```

```{r}
set.seed(1)
cl <- kmeans(vand2[,8:9], 3, nstart = 10)


##png(file="kmeansPlot.png", height=500, width=650)
clusplot(vand2[,c(1:3)], cl$cluster, color=T, lwd=2)
```

```{r}
set.seed(1)
cl <- kmeans(vand2[,8:9], 4, nstart = 10)


##png(file="kmeansPlot.png", height=500, width=650)
clusplot(vand2[,c(1:3)], cl$cluster, color=T, lwd=2)
```



# Pitcher 3



```{r, include=FALSE}

library(dplyr)
vandy_3 <- select(vandy3, RelSpeed, SpinRate, SpinAxis, RelHeight, RelSide, InducedVertBreak, HorzBreak, PlateLocHeight, PlateLocSide)
vand3 <- na.omit(vandy_3)
```


# RelHeight and RelSide with clusters of 2,3, and 4

```{r}
set.seed(1)
cl2 <- kmeans(vand3[,1:3], 2, nstart = 10)


##png(file="kmeansPlot.png", height=500, width=650)
clusplot(vand3[,c(1:3)], cl2$cluster, color=T, lwd=2)
```


```{r}
set.seed(1)
cl2 <- kmeans(vand3[,1:3], 3, nstart = 10)


##png(file="kmeansPlot.png", height=500, width=650)
clusplot(vand3[,c(1:3)], cl2$cluster, color=T, lwd=2)
```


```{r}
set.seed(1)
cl2 <- kmeans(vand3[,1:3], 4, nstart = 10)


##png(file="kmeansPlot.png", height=500, width=650)
clusplot(vand3[,c(1:3)], cl2$cluster, color=T, lwd=2)
```



# RelHeight and RelSide with clusters of 2,3, and 4

```{r}
set.seed(1)
cl2 <- kmeans(vand3[,4:5], 2, nstart = 10)


##png(file="kmeansPlot.png", height=500, width=650)
clusplot(vand3[,c(1:3)], cl2$cluster, color=T, lwd=2)
```


```{r}
set.seed(1)
cl2 <- kmeans(vand3[,4:5], 3, nstart = 10)


##png(file="kmeansPlot.png", height=500, width=650)
clusplot(vand3[,c(1:3)], cl2$cluster, color=T, lwd=2)
```






```{r}
set.seed(1)
cl2 <- kmeans(vand3[,4:5], 4, nstart = 10)


##png(file="kmeansPlot.png", height=500, width=650)
clusplot(vand3[,c(1:3)], cl2$cluster, color=T, lwd=2)
```














# InducedVertBreak and HorzBreak with clusters of 2,3, and 4.


```{r}
set.seed(1)
cl2 <- kmeans(vand3[,6:7], 2, nstart = 10)


##png(file="kmeansPlot.png", height=500, width=650)
clusplot(vand3[,c(1:3)], cl2$cluster, color=T, lwd=2)
```

```{r}
set.seed(1)
cl2 <- kmeans(vand3[,6:7], 3, nstart = 10)


##png(file="kmeansPlot.png", height=500, width=650)
clusplot(vand3[,c(1:3)], cl2$cluster, color=T, lwd=2)
```

```{r}
set.seed(1)
cl2 <- kmeans(vand3[,6:7], 4, nstart = 10)


##png(file="kmeansPlot.png", height=500, width=650)
clusplot(vand3[,c(1:3)], cl2$cluster, color=T, lwd=2)
```


# PlateLocHeight and PlateLocSide with clusters of 2,3, and 4


```{r}
set.seed(1)
cl2 <- kmeans(vand3[,8:9], 2, nstart = 10)


##png(file="kmeansPlot.png", height=500, width=650)
clusplot(vand3[,c(1:3)], cl2$cluster, color=T, lwd=2)
```

```{r}
set.seed(1)
cl2 <- kmeans(vand3[,8:9], 3, nstart = 10)


##png(file="kmeansPlot.png", height=500, width=650)
clusplot(vand3[,c(1:3)], cl2$cluster, color=T, lwd=2)
```

```{r}
set.seed(1)
cl2 <- kmeans(vand3[,8:9], 4, nstart = 10)


##png(file="kmeansPlot.png", height=500, width=650)
clusplot(vand3[,c(1:3)], cl2$cluster, color=T, lwd=2)
```
