---
title: "Vandy"
author: "Josh"
date: "9/16/2019"
output:
  html_document:
    highlight: tango
    theme: lumen
    toc: yes
    toc_depth: 4
---

```{r Setup, include = FALSE}
knitr::opts_chunk$set(echo = FALSE, comment = FALSE, message = FALSE, warning = FALSE)

library(ggplot2)
library(readxl)
library(dplyr)
library(readr)
library(plotly)
library(plyr)
vandy1 <- read_csv("pitcher1_L.csv")

vandy2 <- read_csv("pitcher2_L.csv")
vandy3 <- read_csv("pitcher3_R.csv")
vandy1 <- vandy1[!is.na(vandy1$RelSpeed), ]
vandy2 <- vandy2[!is.na(vandy2$RelSpeed), ]
vandy3 <- vandy3[!is.na(vandy3$RelSpeed), ]
vandy_1 <- vandy1 %>%
  select(TaggedPitchType, PlateLocHeight, PlateLocSide, PitchCall)
vandy_2 <- vandy2 %>%
  select(TaggedPitchType, PlateLocHeight, PlateLocSide, PitchCall)
vandy_3 <- vandy3 %>%
  select(TaggedPitchType, PlateLocHeight, PlateLocSide, PitchCall)
vann <- rbind(vandy_1, vandy_2, vandy_3)
x <- c(-.95,.95,.95,-.95,-.95)
z <- c(1.6,1.6,3.5,3.5,1.6)
sz <- data.frame(x, z)


x <- c(-.95,.95,.95,-.95,-.95)
y <- c(1.6,1.6,3.5,3.5,1.6)
sz <- data.frame(x, y)

x1 <- c(-.95, -.95, 0, .95, .95, -.95)
y1 <- c(0, .15, .25, .15, 0, 0)
hp <- data.frame(x1, y1)

x2 <- c(1.61, 3.11, 3.11, 1.61, 1.61)
y2 <- c(0, 0, 5.5, 5.5, 0)
rhb <- data.frame(x2, y2)

x4 <- c(-1.61, -3.11, -3.11, -1.61, -1.61)
y4 <- c(0, 0, 5.5, 5.5, 0)
lhb <- data.frame(x2, y2)

x3 <- c(2.36)
y3 <- c(2.75)

x5 <- c(-2.36)
y5 <- c(2.75)
cbPalette <- c("#999999", "#E69F00",  "#009E73",  "#0072B2", "#D55E00", "#CC79A7", "#56B4E9", "#F0E442")
```


# Comparing Pitch Velocity

## Calculating Average Pitch Velocity by Pitch Type for Pitcher 1 (Curveball, Fastball, Slider, ChangeUp)

```{r}
vandy1curve <- filter(vandy1, TaggedPitchType == "Curveball")
mean(vandy1curve$RelSpeed)
vandy1fast <- filter(vandy1, TaggedPitchType == "Fastball")
mean(vandy1fast$RelSpeed)
vandy1change <- filter(vandy1, TaggedPitchType == "Slider")
mean(vandy1change$RelSpeed)
vandy1slide <- filter(vandy1, TaggedPitchType == "ChangeUp")
mean(vandy1slide$RelSpeed)
```




## Calculating Average Pitch Velocity by Pitch Type for Pitcher 2

```{r}
vandy2curve <- filter(vandy2, TaggedPitchType == "Curveball")
mean(vandy2curve$RelSpeed)
vandy2fast <- filter(vandy2, TaggedPitchType == "Fastball")
mean(vandy2fast$RelSpeed)
vandy2change <- filter(vandy2, TaggedPitchType == "Slider")
mean(vandy2change$RelSpeed)
vandy2slide <- filter(vandy2, TaggedPitchType == "ChangeUp")
mean(vandy2slide$RelSpeed)
```


## Calculating Average Pitch Velovity by Pitch Type for Pitcher 3

```{r}
vandy3curve <- filter(vandy3, TaggedPitchType == "Curveball")
mean(vandy3curve$RelSpeed)
vandy3fast <- filter(vandy3, TaggedPitchType == "Fastball")
mean(vandy3fast$RelSpeed)
vandy3change <- filter(vandy3, TaggedPitchType == "Slider")
mean(vandy3change$RelSpeed)
vandy3slide <- filter(vandy3, TaggedPitchType == "ChangeUp")
mean(vandy3slide$RelSpeed)
```

Pitcher 1 and Pitcher 2 have very similar pitchspeeds. Pitcher 3 easilly has the highest speed in every sinle pitch.

# Analyzing Strike Zone Pitch Location Plots


## Strike Zone Pitch Location Pitcher 1

```{r, echo=FALSE}
v <- ggplot() +
  geom_path(data = sz, aes(x = x, y = y)) +
  geom_path(data = rhb, aes(x = x2, y = y2), color = "lightgray") +
  geom_path(data = lhb, aes(x = x4, y = y4), color = "lightgray") +
  geom_text(mapping = aes(x = x3, y = y3), label = paste("RHB"), color = "lightgray") +
  geom_text(mapping = aes(x = x5, y = y5), label = paste("LHB"), color = "lightgray") +
  geom_polygon(hp, mapping = aes(x = x1, y = y1), color = "black", fill = "white") +
  coord_equal() +
  ggtitle("Vandy Pitchers by Pitch Type") +
  ##geom_point(data = filter(vandy_2, PitchCall == "BallCalled"),
  geom_point(data = vandy_1,
             aes(x = PlateLocSide, y = PlateLocHeight, color = TaggedPitchType)) +
  scale_color_discrete(limits = c("Fastball" ,"ChangeUp", "Splitter", "Curveball", "Slider", "Sinker")) +
  coord_cartesian(xlim = c(-4.5, 4.5), ylim = c(0, 5.5)) +
  theme(axis.line = element_blank(),
      axis.text.x = element_blank(),
      axis.text.y = element_blank(),
      axis.ticks = element_blank(),
      axis.title.x = element_blank(),
      axis.title.y = element_blank(),
      panel.background = element_blank(),
      panel.border = element_blank(),
      panel.grid.major = element_blank(),
      panel.grid.minor = element_blank(),
      plot.background = element_blank(),
      legend.title = element_blank())
ggplotly(v)
```

## Strike Zone Pitch Location Pitcher 2

```{r, echo = FALSE}
w <- ggplot() +
  geom_path(data = sz, aes(x = x, y = y)) +
  geom_path(data = rhb, aes(x = x2, y = y2), color = "lightgray") +
  geom_path(data = lhb, aes(x = x4, y = y4), color = "lightgray") +
  geom_text(mapping = aes(x = x3, y = y3), label = paste("RHB"), color = "lightgray") +
  geom_text(mapping = aes(x = x5, y = y5), label = paste("LHB"), color = "lightgray") +
  geom_polygon(hp, mapping = aes(x = x1, y = y1), color = "black", fill = "white") +
  coord_equal() +
  ggtitle("Vandy Pitchers by Pitch Type") +
  ##geom_point(data = filter(vandy_2, PitchCall == "BallCalled"),
  geom_point(data = vandy_2,
             aes(x = PlateLocSide, y = PlateLocHeight, color = TaggedPitchType)) +
  scale_color_discrete(limits = c("Fastball" ,"ChangeUp", "Splitter", "Curveball", "Slider", "Sinker")) +
  coord_cartesian(xlim = c(-4.5, 4.5), ylim = c(0, 5.5)) +
  theme(axis.line = element_blank(),
      axis.text.x = element_blank(),
      axis.text.y = element_blank(),
      axis.ticks = element_blank(),
      axis.title.x = element_blank(),
      axis.title.y = element_blank(),
      panel.background = element_blank(),
      panel.border = element_blank(),
      panel.grid.major = element_blank(),
      panel.grid.minor = element_blank(),
      plot.background = element_blank(),
      legend.title = element_blank())
ggplotly(w)
```

## Strike Zone Pitch Location Pitcher 3

```{r}
u <- ggplot() +
  geom_path(data = sz, aes(x = x, y = y)) +
  geom_path(data = rhb, aes(x = x2, y = y2), color = "lightgray") +
  geom_path(data = lhb, aes(x = x4, y = y4), color = "lightgray") +
  geom_text(mapping = aes(x = x3, y = y3), label = paste("RHB"), color = "lightgray") +
  geom_text(mapping = aes(x = x5, y = y5), label = paste("LHB"), color = "lightgray") +
  geom_polygon(hp, mapping = aes(x = x1, y = y1), color = "black", fill = "white") +
  coord_equal() +
  ggtitle("Vandy Pitchers by Pitch Type") +
  ##geom_point(data = filter(vandy_2, PitchCall == "BallCalled"),
  geom_point(data = vandy_3,
             aes(x = PlateLocSide, y = PlateLocHeight, color = TaggedPitchType)) +
  scale_color_discrete(limits = c("Fastball" ,"ChangeUp", "Splitter", "Curveball", "Slider", "Sinker")) +
  coord_cartesian(xlim = c(-4.5, 4.5), ylim = c(0, 5.5)) +
  theme(axis.line = element_blank(),
      axis.text.x = element_blank(),
      axis.text.y = element_blank(),
      axis.ticks = element_blank(),
      axis.title.x = element_blank(),
      axis.title.y = element_blank(),
      panel.background = element_blank(),
      panel.border = element_blank(),
      panel.grid.major = element_blank(),
      panel.grid.minor = element_blank(),
      plot.background = element_blank(),
      legend.title = element_blank())
ggplotly(u)
```


One thing we noticed when analyzing fastballs was that pitcher 2 and 3 were in the strike zone while pitcher 1 was a bit less accurate. Pitcher 1 throws the most change ups by far and seems to consistently in the strike sone. As far as the curveball goes pitcher 2 and 3 are pretty conistent on placement while pitcher is very off. All of this tells us that pitcher 2 and 3 are more consistent while pitcher 1 is not.



```{r}
ggplot(vandy1, aes(x = PitchCall)) +
geom_bar(fill = "blue", color = "orange") + facet_wrap(~ TaggedPitchType) +
scale_x_discrete(label=abbreviate) +
  theme_dark()
```

```{r}
ggplot(vandy2, aes(x = PitchCall)) +
geom_bar(fill = "blue", color = "orange") + facet_wrap(~ TaggedPitchType) +
scale_x_discrete(label=abbreviate) +
  theme_dark()
```



```{r}
ggplot(vandy3, aes(x = PitchCall)) +
geom_bar(fill = "blue", color = "orange") + facet_wrap(~ TaggedPitchType) +
scale_x_discrete(label=abbreviate) +
  theme_dark()
```


From these bar graphs we can tell that pitcher 2 tends to throw more strikes than balls. This is not the case with pitcher 1 and 3. Pitcher 3 seems to have the most variety in his game as the other 2 don't throw many curveballs or change-ups. Pitcher 2 also has the least ammount of balls in play so we have to give pither 2 the edge on this one. In conclusion pitcher 2 and 3 seem to be more skilled than pitcher 1. Pitcher 3 is more versatile as a pitcher but pitcher 2 has better outcomes.
