---
title: "Untitled"
author: "Jefferson Ong"
date: "1/30/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


presentation layout: 
slide1: present dataset(showing head), 

slide 2: look at variables and explain them, 
slide : picture for index, 


slide 3: month, scatterplot faceted,


slide 4: temp, scatterplot, 
slide : Lm(explain corr), of temp


slide 7: Rain, scatterplot(maybenot)

slide 8: FFMC, scatter

slide 9: DC, linear






```{r}
library(tidyverse)
library(readxl)
forest <- read_csv("forestfires.csv")
head(forest)


```

## Forest Fire Dataset

https://archive.ics.uci.edu/ml/datasets/Forest+Fires

This is a difficult regression task, where the aim is to predict the burned area of forest fires, in the northeast region of Portugal, by using meteorological and other data 


Data Set Information:

In [Cortez and Morais, 2007], the output 'area' was first transformed with a ln(x+1) function.
Then, several Data Mining methods were applied. After fitting the models, the outputs were
post-processed with the inverse of the ln(x+1) transform. Four different input setups were
used. The experiments were conducted using a 10-fold (cross-validation) x 30 runs. Two
regression metrics were measured: MAD and RMSE. A Gaussian support vector machine (SVM) fed
with only 4 direct weather conditions (temp, RH, wind and rain) obtained the best MAD value:
12.71 +- 0.01 (mean and confidence interval within 95% using a t-student distribution). The
best RMSE was attained by the naive mean predictor. An analysis to the regression error curve
(REC) shows that the SVM model predicts more examples within a lower admitted error. In effect,
the SVM model predicts better small fires, which are the majority.

Attribute Information:

For more information, read [Cortez and Morais, 2007].
1. X - x-axis spatial coordinate within the Montesinho park map: 1 to 9
2. Y - y-axis spatial coordinate within the Montesinho park map: 2 to 9
3. *month* - month of the year: 'jan' to 'dec'
4. day - day of the week: 'mon' to 'sun'
5. *FFMC* - FFMC index from the FWI system: 18.7 to 96.20
6. *DMC* - DMC index from the FWI system: 1.1 to 291.3
7. *DC* - DC index from the FWI system: 7.9 to 860.6
8. *ISI* - ISI index from the FWI system: 0.0 to 56.10
9. *temp* - temperature in Celsius degrees: 2.2 to 33.30
10. RH - relative humidity in %: 15.0 to 100
11. wind - wind speed in km/h: 0.40 to 9.40
12. *rain* - outside rain in mm/m2 : 0.0 to 6.4
13. area - the burned area of the forest (in ha): 0.00 to 1090.84
(this output variable is very skewed towards 0.0, thus it may make
sense to model with the logarithm transform).




IMAGE FOR POWERPOINT _ https://www.nwcg.gov/sites/default/files/publications/images/437-cffdrs-overview-fwi-flowchart.png 
```{r}
firstregress <- lm(area ~ X + Y + month + day + FFMC + DMC + DC + ISI + temp + RH + wind + rain, forest)
summary(firstregress)
```


```{r}
secondregress <- lm(area~DC + DMC, forest)
summary(secondregress)
```

```{r}
#thirdregress <- lm(area~, forest)
#summary(thirdregress)
```

```{r}
str(forest)
```

```{r}
cor(forest$FFMC + forest$, forest$area)
```



```{r}
str(forest)
```



```{r}
pairs( ~  area + X + Y + FFMC + DMC + DC + ISI + temp + RH + wind + rain,data = forest)


#cor.test( ~ Zsagarin + Fravg,
#         data=data,
#         method = "pearson",
#         conf.level = 0.95)
```


```{r}
fourthregress  <- lm(area ~ RH + rain  + temp + wind , forest)
summary(fourthregress)

fifthregress  <- lm(area ~  RH , forest)
summary(fifthregress)


# VVVVV pvalue is .0261 with just temp
sixthregressA <- lm(area ~ rain, forest)
summary(sixthregressA)



seventhregress <-  lm(area ~ log(DC), forest)
summary(seventhregress)

one <- lm(area ~  RH, forest)
summary(one)

ggplot(forest, aes(x = RH, y = area))+
  geom_point() +
  geom_smooth(method = "lm")
  
```

```{r}
one <- lm(area ~ temp + RH, forest)
summary(one)

sixthregressA <- lm(area ~ temp, forest)
summary(sixthregressA)

```




```{r}
library(ggplot2)
ggplot(data = orest, aes(y = area, x = wind))+
  geom_point()+
  geom_smooth(method = "lm")


```



```{r}
library(remotes)
rm.outlier(forest, temp, within_subjects = FALSE, wid = NULL,
  median = FALSE, sd_times = 3, print = FALSE)
```

```{r}
ggplot(forest, aes(y = area, x = temp)) +
  geom_point() +
  geom_smooth(method = lm) +
  theme_bw()

ggplot(forest, aes(y = area, x = wind)) +
  geom_point() +
  geom_smooth(method = lm) +
  theme_bw()
ggplot(forest, aes(y = area, x = temp+ RH)) +
  geom_point() +
  geom_smooth(method = lm) +
  theme_bw()
ggplot(forest, aes(y = area, x = RH)) +
  facet_wrap(.~forest$month) +
  geom_boxplot()
#  geom_smooth(method = lm) +
#  theme_bw()

# Jan - no fire because snowy
# Feb - few fires because less snow fall
# Mar - not wide spread of fire because of rain
# April
# May
# June
# July
# Aug - 
# Sep
# Oct
# Nov
# Dec




#Summer in Portugal sits between June and mid-September while the winter season falls between December and February. While Portugal does have four distinct seasons, the respective conditions of summer and winter do tend to linger on in autumn and spring
```









